fGFPCA dynamic prediction

### 2023-02-20 f-GFPCA.Rmd
clean scripts

```{r, laplace_ped}
# clean MLE out-of-sample prediction results
out_pred_clean_laplace <- function(subjID=1, mid=mid){
  # Prediction
  pred_t195 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=195) %>% select(-eta_i))
  pred_t395 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=395) %>% select(-eta_i))
  pred_t595 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=595) %>% select(-eta_i))

@@

```{r, fig.width=12, fig.height=4}
# compare point prediction with in-sample
# subject 1
grid.arrange(
  out_pred_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample MLE"),

@@

  out_pred_lap_subj3 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 3 out-of-sample Laplace"),

@@

  out_pred_lap_subj4 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 4 out-of-sample Laplace"),

@@

### In-/Out-of-sample PC scores

>>>>

```{r, laplace_pred}
# clean MLE out-of-sample prediction results
out_pred_clean_laplace <- function(subjID=1, mid=mid){
  # Prediction
  pred_t195 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=195) %>% select(-eta_i))
  pred_t395 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=395) %>% select(-eta_i))
  pred_t595 <- out_pred_laplace(fpca_fit = fpca_fit, df_new = df %>% filter(id==subjID & sind_inx<=595) %>% select(-eta_i))

@@

```{r, fig.width=8, fig.height=4}
# compare point prediction with in-sample
# subject 1
grid.arrange(
  out_pred_subj1 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 1 out-of-sample MLE"),

@@

  out_pred_lap_subj3 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 2 out-of-sample Laplace"),

@@

  out_pred_lap_subj4 %>% 
    group_by("type") %>%
    ggplot()+
    geom_line(aes(x=mid, y=true), col = "black")+
    geom_line(aes(x=mid, y=eta, col= type, group = type), linetype="dashed", na.rm = T,
              show.legend = FALSE)+
    labs(x="Binned time", title = "Subject 2 out-of-sample Laplace"),

@@

## Predictive performance measures

#### Measures

- Focusing on prediction of individual tracks
- Integrated error: for latent process

$$\frac{1}{N}\sum_{i=1}^N\sum_s(\hat{\eta}_{is}-\eta_{is})^2$$
where s covers only unobserved tracks

- Integrated AUC: for binary outcome
- Perhaps, also computation time?

<span style="color:red;">The problem with measures above is that predictions are made on a different grid than observations. I can think of two potential ways to address this issue: 1) only evaluate at bin midpoints; 2) use interpolation to extend latend function from binned grid to the original grid (in GFPCA paper this is done through projection. ) </span>

#### Reference methods

Also, we need to compare with a existing method for dynamic prediction of binary outcomes. Ideals, there should be a FDA component to this methods.

1. registr::bfpca:

This method is not a method for prediction. If we would like to use it for binary prediction, the two methods will only differ in the first half (FPCA). But this comparision has already been made in the GFPCA paper.

This method also has convergence issue.

2. Functional mixed models (Intercept-only)

Can these methods handle binary function? They are also very, very slow to fit on the full dataset.

3. GLMM
Also had convergence issues.

## In-/Out-of-sample PC scores
